### 2015-05-13 related.tex
update

Statically analyzing concurrent programs to detect races is a promising alternative to dynamic analyzers, which commonly face code coverage issues as they rely on a (controlled) scheduler for exploring execution paths~\cite{musuvathi2008finding}. Warlock~\cite{sterling1993warlock} and LockLint~\cite{oracle2010locklint} are notable examples of static race analyzers. However, both tools have limited applicability, as they heavily rely on user annotations. \textsc{Whoop}, on the other hand, does not require any source code modifications, and thus can be applied with minimal effort.

Most related to our static lockset analysis are the static lockset analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have significant limitations. The authors of RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To limit these false positives, RELAY employs post-analysis unsound filters, but these can also filter out true races. Although Locksmith successfully detected data races in several small Pthreads applications and 7 medium-sized Linux device drivers, it also reported a significant number of false positives. The authors also reported that Locksmith was unable to run on several large programs, showcasing its limited scalability. In contrast, \whoop is based on novel over-approximation techniques and uses modern SMT solvers to accelerate state-of-the-art concurrency bug-finders, such as \corral, for precise \emph{and} scalable analysis.

>>>>

Statically analyzing concurrent programs for races is a promising alternative to dynamic analyzers, which typically rely on a (controlled) scheduler for exploring execution paths, and thus can face code coverage issues~\cite{musuvathi2008finding}. Warlock~\cite{sterling1993warlock} and LockLint~\cite{oracle2010locklint} are notable examples of static race analyzers. However, both tools have limited applicability, as they heavily rely on user annotations. \textsc{Whoop}, on the other hand, does not require any source code modifications, and thus can be applied with minimal effort.

Most related to our lockset analysis are the static lockset analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have several limitations. RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To tackle this issue, RELAY employs post-analysis unsound filters, but these can also filter out real bugs. Locksmith was successfully applied in several small Pthreads applications and 7 medium-sized Linux device drivers, but the authors reported that the tool was unable to run on several large programs, showcasing its limited scalability. \whoop aims to achieve scalability \emph{and} precision: the first, via novel symbolic pairwise lockset analysis; and the second, by accelerating \corral, an industrial-strength precise bug-finder.
